---
title: "Week7_Lab1:Map making"
author: "Shanelle A. Wikramanayake"
date: "3/8/2021"
output:
  html_document:
    toc: TRUE #create table of contents
    toc_float: TRUE #To the left
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path = "../Outputs/")
```

# Introduction  
Today we are going to work with data to learn _map making in R_ 

# Load Libraries
```{r, warning= FALSE, message=FALSE}

library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(kableExtra)
library(tidytuesdayR)
```

# Load and inspect data 
```{r}
tuesdata <- tidytuesdayR::tt_load('2021-01-26')
plastics <- tuesdata$plastics
view(plastics)

plastics_data <- plastics %>% 
  select(country,pvc, hdpe, ldpe, pet, pp) %>% 
  group_by(country) %>% 
  summarise(mean_pvc = mean(pvc), 
            mean_hdpe = mean(hdpe),
            mean_ldpe = mean(ldpe), 
            mean_pet = mean(pet), 
            mean_pp = mean(pp)) 
plastics_data %>% 
  kbl()


```

# Plastic pollution across the globe
Lets visualize plastic pollution across the globe
First, lets look at the distribution of mean PVC plastics around the world.
```{r}
world <- map_data("world") #creates a df of polygons of the world

plastics_world <- plastics %>% 
  select(country,pvc, hdpe, ldpe, pet, pp) %>% 
  group_by(country) %>% 
  summarise(mean_pvc = mean(pvc), 
            mean_hdpe = mean(hdpe),
            mean_ldpe = mean(ldpe), 
            mean_pet = mean(pet), 
            mean_pp = mean(pp)) %>% 
  rename(region = country)
view(plastics_world)

plastics_world <- plastics_world %>% 
  inner_join(world, 
             by = "region")
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               colour = "black")+
  geom_polygon(data = plastics_world, 
               aes(x = long, 
                   y = lat,
                   group = group,
                   fill = mean_pvc))+
  scale_fill_viridis_c()+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "light blue"))+
  coord_map(projection = "mercator", 
            xlim = c(-180, 180))+
  labs(title = "Mean PVC Plastic Count for Different Countries", 
       x = "Longitude", 
       y = "Latitude", 
       fill = "Mean PVC count")
  
 

```
Now, lets look at the distribution of mean **High density polyethylene** (HDPE) plastics around the world.
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               colour = "black")+
  geom_polygon(data = plastics_world, 
               aes(x = long, 
                   y = lat,
                   group = group,
                   fill = mean_hdpe))+
  scale_fill_viridis_c()+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "light blue"))+
  coord_map(projection = "mercator", 
            xlim = c(-180, 180))+
  labs(title = "Mean HDPE Plastic Count for Different Countries", 
       x = "Longitude", 
       y = "Latitude", 
       fill = "Mean HDPE count")
  
```

Now, lets look at the distribution of mean **Low density polyethylene** (LDPE) plastics around the world.
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               colour = "black")+
  geom_polygon(data = plastics_world, 
               aes(x = long, 
                   y = lat,
                   group = group,
                   fill = mean_ldpe))+
  scale_fill_viridis_c()+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "light blue"))+
  coord_map(projection = "mercator", 
            xlim = c(-180, 180))+
  labs(title = "Mean LDPE Plastic Count for Different Countries", 
       x = "Longitude", 
       y = "Latitude", 
       fill = "Mean LDPE count")
```
 
Now, lets look at the distribution of mean **Polyester** (PET) plastics around the world.
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               colour = "black")+
  geom_polygon(data = plastics_world, 
               aes(x = long, 
                   y = lat,
                   group = group,
                   fill = mean_pet))+
  scale_fill_viridis_c()+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "light blue"))+
  coord_map(projection = "mercator", 
            xlim = c(-180, 180))+
  labs(title = "Mean PET Plastic Count for Different Countries", 
       x = "Longitude", 
       y = "Latitude", 
       fill = "Mean PET count")

```


